The Decline in U.S Birthrates in Recent Years is Indicative of Cultural and Economic Changes

### 2022-02-28 Output/paper 2.Rmd
Final changes

---
title: "The Decline in U.S Birthrates in Recent Years is Indicative of Socio-economic Changes"
author: "Olaedo Okpareke, Arsh Lakhanpal, Swarnadeep Chattopadhyay"
thanks: 'Code, data and replication are available at: https://github.com/OlaedoOkpareke/Paper2Repository and https://doi.org/10.48152/ssrp-8bs7-7e32' 
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Birth rates provide useful information on population growth, and an above-replacement birth rate is indicative of stability. However, previous research has found that birth rates in industrialized societies such as the U.S have been decreasing in recent years. We use data from the American Economic Association to analyze this by using graphs and tables to observe the trend in birth rates of different demographics, using R [@citeR] and other packages. We also observe the trends behind changing economic and cultural factors that could affect the birth rates. We find that birth rates of young women of different races have been on a steep decline over the past 20 years, and factors that are known to decrease birth rates have increased. This report shows that the decline in birth rates is the result of cultural and economic changes between two generations of women. While this indicates that the U.S has become more culturally liberated, the declining birth rates showcases systemic effects of the worsening economy and increasing costs on the average American woman."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references2.bib

@@

## Results

@@

White birth rates have decreased significantly, and the 15-24 age group contributes more than 18% of declining birth rates. This decrease in white birth rates is a well known talking point of the far right, which is used to harass people of color, especially mothers. This concern over the white birth rate decline (often coded as 'True Americans') and the notion of 'being replaced' has led to harmful anti-immigration policies, as well as the rise of far right hate groups throughout the country [@Chermak2013].The increase and proliferation of far right ideologies that white birth rates are decreasing leading to a less white country are fueled in part by the way that these statistics can be shown; without the proper context to show why they are decreasing. Many pro-life advocates have also noted declining birth rates as a major reason for their antipathy towards abortion, which has caused an increase in attacks on planned parenthood clinics as they are believed to be a symptom of societal decay. [@Primrose].

@@

# State Difference

@@

There were datasets in which we had to generalize about a specific population based on data of a general population. For example, in (Figure \@ref(fig:figseven)), the map data was composed of responses taken from women aged from 20-34, as information for the 15 to 24 age group was not available. Due to this, conclusions that are made from this map are likely not accurate as they are not the true statistics of the 15 to 24 age group. Our research question wanted to focus on the trend of birth rates over a 20 year period, specifically from 2000 to 2020. However, for several datasets, only data up to 2019 was available. Due to this, instead of being able to compare 2000 and 2020 explicitly, we had to generalize trends that we see in 2019 to be at least similar to trends that we would see in 2020. We had the same problem with (Table \@ref(tab:tabtwo)), in which we had to approximate 2004 data for 2000 trends, even though situations would be different. 
In addition to this, the year ranges for millennials and generation Z are subjective, as many different years have been suggested as the start and end for these two generations. Thus the years chosen for this report are completely arbitrary.

>>>>

---
title: "The Decline in U.S Birthrates in Recent Years is Indicative of Cultural and Economic Changes"
author: "Olaedo Okpareke, Arsh Lakhanpal, Swarnadeep Chattopadhyay"
thanks: 'Code, data and reproduction package are available at: https://github.com/OlaedoOkpareke/Paper2Repository and https://doi.org/10.48152/ssrp-8bs7-7e32' 
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Birth rates provide useful information on population growth, and an above-replacement birth rate is indicative of stability. However, previous research has found that birth rates in industrialized societies such as the U.S have been decreasing in recent years. We use data from the American Economic Association to analyze this by using graphs and tables to observe the trend in birth rates of different demographics, using R [@citeR] and other packages. We also observe the trends behind changing economic and cultural factors that could affect the birth rates. We find that birth rates of young women of different races have been on a steep decline over the past 20 years, and factors that are known to decrease birth rates have increased. This report shows that the decline in birth rates is the result of cultural and economic changes between two generations of women. While this indicates that the U.S has become more culturally liberated, the declining birth rates showcases systemic effects of the worsening economy and increasing costs on the average American woman."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references2.bib

@@

# Results

@@

White birth rates have decreased significantly, and the 15-24 age group contributes more than 18% of declining birth rates. This decrease in white birth rates is a well known talking point of the far right, which is used to harass people of color, especially mothers. This concern over the white birth rate decline (often coded as 'True Americans') and the notion of 'being replaced' has led to harmful anti-immigration policies, as well as the rise of far right hate groups throughout the country [@Chermak2013].The increase and proliferation of far right ideologies that white birth rates are decreasing leading to a less white country are fueled in part by the way that these statistics can be shown; without the proper context to show why they are decreasing. Many pro-life advocates have also noted declining birth rates as a major reason for their antipathy towards abortion, which has caused an increase in attacks on planned parenthood clinics as they are believed to be a symptom of societal decay [@Primrose].

@@

## State Differences

@@

There were datasets in which we had to generalize about a specific population based on data of a general population. For example, in (Figure \@ref(fig:figseven)), the map data was composed of responses taken from women aged from 20 to 34, as information for the 15 to 24 age group was not available. Due to this, conclusions that are made from this map are likely not accurate as they are not the true statistics of the 15 to 24 age group. Our research question wanted to focus on the trend of birth rates over a 20 year period, specifically from 2000 to 2020. However, for several datasets, only data up to 2019 was available. Due to this, instead of being able to compare 2000 and 2020 explicitly, we had to generalize trends that we see in 2019 to be at least similar to trends that we would see in 2020. We had the same problem with (Table \@ref(tab:tabtwo)), in which we had to approximate 2004 data for 2000 trends, even though situations would be different. 
In addition to this, the year ranges for millennials and generation Z are subjective, as many different years have been suggested as the start and end for these two generations. Thus the years chosen for this report are completely arbitrary.
